FictionEro2 - Determinants of Reality Beliefs

### 2025-05-08 study2/analysis/2_emotion.qmd
data analysis
lacking final figures and rendering emotions analysis script

**notes**: How about confidence on all AI and all photos.

@@

df <- read.csv("../data/data.csv") |> 
  right_join(
    select(dfsub, Participant, Mobile, starts_with(c("Feedback_","BAIT")), COPS_Frequency, SexualActivity,- Feedback_Text),
    by = "Participant"
    ) |> 
  mutate(Condition = case_when(
    Condition == "Fiction" ~ "AI-Generated",
    Condition == "Reality
    " ~ "Photograph"
  )) |>
  mutate(Condition = fct_relevel(Condition, "Photograph", "AI-Generated"),
         Gender =  fct_relevel(Gender, "Male", "Female"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency)),
         SexualActivity_num = as.numeric(as.factor(SexualActivity)),
         ConditionBelief = case_when(
           Condition == "Photograph" & Realness > 0 ~ "True",
           Condition == "AI-Generated" & Realness < 0 ~ "True",
           .default = "False",
         )) |>
  rename(AllRealConfidence = "Feedback_AllRealConfidence",
         AllFakeConfidence = "Feedback_AllFakeConfidence",
         Enjoyment = "Feedback_Enjoyment"
         )|>
  mutate(across(starts_with("Feedback_"), as.factor))

@@

## re-scaled variables
scale_values <- function(x){(x-min(x))/(max(x)-min(x))}

df$Arousal <- scale_values(df$Arousal)
df$Enticing <- scale_values(df$Enticing)
df$Valence <- scale_values(df$Valence)
```

@@

## Manipulation Check

@@

To summarise, women rate stimuli on average significantly lower in arousal, enticiment and valence than men.

@@

Arousal ratings were relatively well-distributed across the scale, with the highest frequency for "Not at all" and a gradual decrease across higher ratings.

The pattern of arousal ratings was similar regardless of whether participants believed the images were AI-generated or real. Among male participants, arousal ratings were relatively evenly distributed but tended to cluster around the middle of the scale, independent of their belief about the images' origin. Female participants, on the other hand, showed a pronounced peak at the lowest arousal rating, irrespective of their perception of the images as AI-generated. Participants with other gender identities provided fewer responses overall, but their arousal ratings generally leaned toward moderate levels, particularly in the "Not Believed" condition.

```{r}
arousal <- df |> 
  ggplot(aes(x=Arousal)) +
  geom_bar(fill="#03A9F4") +
  scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Arousal", ) +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

@@

df |> 
  mutate(ConditionBelief = as.factor(ConditionBelief)) |> 
  mutate(ConditionBelief = case_when(
    ConditionBelief == "False" ~ "Not Believed",
    ConditionBelief == "True" ~ "Believed",)) |>
  ggplot(aes(x=Arousal, fill=Gender)) +
  geom_bar(position = "identity") +
  facet_grid(Gender~ConditionBelief, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Female"="#E91E63", "Male"="#FF9800"), guide="none") +
  # scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Arousal to images believed or not to be AI-Gnerated") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

**Note**:  Model m3, which includes random intercepts and random slopes for Gender at the stimulus level, exhibits relatively high correlations between the random intercepts and slopes. This suggests that the random effects are strongly correlated, which could indicate overfitting.

@@

test_bf(m1,m2, m3)
compare_performance(m1, m2)

# the model with participants only as random effecs seems to be the best
```

@@

Enticement ratings were relatively well-distributed across the scale.

Male participants predominantly provided responses in the mid-to-high enticement range across both conditions. Female participants showed a pronounced peak at the lowest enticement rating when they did not believe the image was AI-generated, with a similar but less pronounced pattern in the believed condition. Participants identifying as 'Other' responses were more concentrated around the middle of the scale.

```{r}
enticiment <- df |> 
  ggplot(aes(x=Enticing)) +
  geom_bar(fill="#03A9F4") +
  scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Enticement", ) +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

@@

df |> 
  mutate(ConditionBelief = as.factor(ConditionBelief)) |> 
  mutate(ConditionBelief = case_when(
    ConditionBelief == "False" ~ "Not Believed",
    ConditionBelief == "True" ~ "Believed",)) |>
  ggplot(aes(x=Enticing, fill=Gender)) +
  geom_bar(position = "identity", binwidth = 0.5) +
  facet_grid(Gender~ConditionBelief, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Female"="#E91E63", "Male"="#FF9800"), guide="none") +
  # scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Enticiment to images believed or not to be AI-Gnerated") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

@@

```{r}
estimate_relation(m_e) |> 
  ggplot(aes(x=Gender, y=Predicted)) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color = Gender), position=position_dodge(width=0.5)) +
 scale_color_manual(values=c("Male"="#E91E63", "Female"="#FF9800"), guide="none") +  labs(y="Arousal") +
  theme_abyss()
```

@@

Valence ratings were concetrated more around the mid-to-high valence.

Participants showed the same pattern of results regardless of whether they believed the images to be Ai-generated or not. 
Males tended to to have the majority of their responses concentrated in the mid-to-high valence range. Female majoity valence ratings were on on the lower (0) and midle end of the scale (3/4).

```{r}
valence <- df |> 
  ggplot(aes(x=Valence)) +
  geom_bar(fill="#03A9F4") +
  scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Valence", ) +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

@@

df |> 
  mutate(ConditionBelief = as.factor(ConditionBelief)) |> 
  mutate(ConditionBelief = case_when(
    ConditionBelief == "False" ~ "Not Believed",
    ConditionBelief == "True" ~ "Believed",)) |>
  ggplot(aes(x=Valence, fill=Gender)) +
  geom_bar(position = "identity", binwidth = 0.5) +
  facet_grid(Gender~ConditionBelief, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Female"="#E91E63", "Male"="#FF9800"), guide="none") +
  # scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Valence to images believed or not to be AI-Gnerated") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

@@

<!-- Key Findings -->

@@

```{r}
estimate_relation(m_v) |> 
  ggplot(aes(x=Gender, y=Predicted)) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color = Gender), position=position_dodge(width=0.5)) +
 scale_color_manual(values=c("Male"="#E91E63", "Female"="#FF9800"), guide="none") +  labs(y="Arousal") +
  theme_abyss()
```

@@

marginaleffects::avg_slopes(m_arousal, variables="Condition", by="Gender") 
#   results_table()

@@

The analysis revealed a significant interaction effect only for **positive attitudes towards AI** among females. Specifically, females with higher positive attitudes towards AI rated AI-generated images as more arousing, relative to when they viewed photographs

@@

```{r}
m <- glmmTMB::glmmTMB(Arousal ~ Gender / (Condition * SexualActivity_num) + (Condition | Participant) + (1|Stimulus),
                      data=filter(df, ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

@@

This model examines how Gender and Condition influence enticement ratings, treating participant and items as random intercepts.

@@

**Note**. 'Other' was removed due to the very small percentage of individuals identifying as such.

@@

Looking at **feedback as a moderator**, the results showed that enticement ratings for images labelled as AI-Generated were moderated by the feedback that AI is less Attractive for males, and that difference between the AI-Generated images and real photos is obvious for females.

@@

           "Feedback_DiffObvious", "Feedback_DiffNone",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing",
           "Feedback_AIMoreAttractive", "Feedback_LessArousing",
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  moderator_believed <- rbind(moderator_believed,
                             test_moderator(filter(df, ConditionBelief == "True"), "Enticing", m))
}

@@

This model examines the effect of Gender and Condition on Valence ratings, while accounting for random variability due to participants and items (i.e., random intercepts).

@@

```{r}
for(m in c("Feedback_DiffObvious", "Feedback_DiffNone","Feedback_DiffSubtle",

           "Feedback_AIMoreArousing", "Feedback_AILessArousing",
           "Feedback_AIMoreAttractive", "Feedback_LessArousing",

           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed",
           "Feedback_AllReal", "Feedback_AllFake")) {
  moderator_believed <- rbind(moderator_believed,
                             test_moderator(filter(df, ConditionBelief == "True"), "Valence", m))
}

@@

```{r}
df$Emotionality <- normalize(predict(pca)$PC1)

emotionality <- df |> 
  mutate(Emotionality_Extreme = ifelse(Emotionality %in% c(0, 1), "Extreme", "Not extreme")) |>
  ggplot(aes(x=Emotionality, fill = Gender)) +
  geom_histogram(aes(color=Emotionality_Extreme), position = "identity", bins=80) +
  facet_grid(~Gender, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Female"="#48c9b0", "Male"="#FF9800"), guide="none") +
  scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Emotionality") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

@@

This model looks at the effect of gender and condition on emotionality, accounting for random variability due to participants and items (i.e., random intercepts).

@@

```{r}
m <- glmmTMB::glmmTMB(Emotionality ~ Gender /Condition + (1| Participant) + (1|Stimulus),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

@@

Looking at **feedback** as a moderator, the results show that for males, emotionality ratings for AI-Generated images were negatively moderated by the feedback that AI images were more attractive.

@@

```{r}
for(m in c("Feedback_DiffObvious", "Feedback_DiffNone","Feedback_DiffSubtle",

           "Feedback_AIMoreArousing", "Feedback_AILessArousing",
           "Feedback_AIMoreAttractive", "Feedback_LessArousing",

           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed",
           "Feedback_AllReal", "Feedback_AllFake")) {
  moderator_believed <- rbind(moderator_believed,
                             test_moderator(filter(df, ConditionBelief == "True"), "Emotionality", m))
}

@@

-

@@

>>>>

@@

df <- read.csv("../data/data.csv") |> 
  right_join(
    select(dfsub, Participant, Mobile, starts_with(c("Feedback_","BAIT")), COPS_Frequency, SexualActivity,- Feedback_Text),
    by = "Participant"
    ) |> 
    datawizard::rescale(select= c("Arousal", "Enticing", "Valence"), range=c(0, 6), to=c(0,1)) |> 
  datawizard::rescale(select= c("Realness"), range=c(-3,3), to=c(0,1)) |>
  mutate(Condition = case_when(
    Condition == "Fiction" ~ "AI-Generated",
    Condition == "Reality" ~ "Photograph"
  )) |>
  mutate(Condition = fct_relevel(Condition, "Photograph", "AI-Generated"),
         Gender =  fct_relevel(Gender, "Male", "Female"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency)),
         SexualActivity_num = as.numeric(as.factor(SexualActivity)),
         ConditionBelief = case_when(
           Condition == "Photograph" & Realness > 0.5 ~ "True",
           Condition == "AI-Generated" & Realness < 0.5 ~ "True",
           .default = "False",
         )) |>
  rename(AllRealConfidence = "Feedback_AllRealConfidence",
         AllFakeConfidence = "Feedback_AllFakeConfidence",
         Enjoyment = "Feedback_Enjoyment"
         )|>
  mutate(across(starts_with("Feedback_"), as.factor))

@@

@@

## Visualisation of Key Variables

@@

To summarise, **women** rate stimuli on average **significantly lower** in arousal, enticiment and valence than **men**.

@@

In general, males exhibited greater variability in their arousal ratings, while females tended to provide more lower arousal ratings.

The pattern of arousal ratings for images that participants believed to be either AI-generated or photos (i.e., initially labeled as photos and later rated as photos, or initially labeled as AI and later rated as AI), as well as for images where participants disagreed with the original labels, was generally similar across genders.

```{r}
arousal <- df |> 
  ggplot(aes(x=Arousal, fill = Gender)) +
  geom_bar() +
  scale_y_sqrt(expand = c(0, 0)) +
  scale_x_continuous(
    breaks = c(0, 1),
    labels = c("Not at all","Very much")) +
  labs(title = "Arousal", ) +
  facet_wrap(~Gender) +
   scale_fill_manual(values=c("Male"="#E91E63", "Female"="#FF9800"), guide="none") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

@@

df |> 
  mutate(ConditionBelief = as.factor(ConditionBelief)) |> 
  mutate(ConditionBelief = case_when(
    ConditionBelief == "False" ~ "Not Believed",
    ConditionBelief == "True" ~ "Believed",)) |>
  ggplot(aes(x=Arousal, fill=Gender)) +
  geom_bar(position = "identity") +
   scale_x_continuous(
    breaks = c(0, 1),
    labels = c("Not at all","Very much")) +
  facet_grid(Gender~ConditionBelief, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Male"="#E91E63", "Female"="#FF9800"), guide="none") +
  # scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Arousal Ratings for Images Based on Belief in Original Labels") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

@@

test_bf(m1,m2, m3)
compare_performance(m1, m2, m3)

# Model m3, which includes random intercepts for Participants and both random intercepts and slopes for Gender at the Stimulus level, seems to be the best
```

@@

In general, males tended to rate images higher in enticement more often than females.

Enticing ratings for images where participants believed in the original labels, as well as for those where they disagreed with the labels, were similar for males.

For females, although a similar pattern was observed, they tended to provide more lower enticement ratings, particularly for images where they believed in the original labels.

```{r}
enticiment <- df |> 
  ggplot(aes(x=Enticing, fill = Gender)) +
  geom_bar() +
  scale_y_sqrt(expand = c(0, 0)) +
  scale_x_continuous(
    breaks = c(0, 1),
    labels = c("Not at all","Very much")) +
  labs(title = "Enticement", ) +
  facet_wrap(~Gender) +
   scale_fill_manual(values=c("Male"="#E91E63", "Female"="#FF9800"), guide="none") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

@@

df |> 
  mutate(ConditionBelief = as.factor(ConditionBelief)) |> 
  mutate(ConditionBelief = case_when(
    ConditionBelief == "False" ~ "Not Believed",
    ConditionBelief == "True" ~ "Believed",)) |>
  ggplot(aes(x=Enticing, fill=Gender)) +
  geom_bar(position = "identity", binwidth = 0.5) +
  scale_x_continuous(
    breaks = c(0, 1),
    labels = c("Not at all","Very much")) +
  facet_grid(Gender~ConditionBelief, scales="free_y") +
  scale_fill_manual(values=c("Female"="#E91E63", "Male"="#FF9800"), guide="none") +
  # scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Enticement Ratings for Images Based on Belief in Original Labels") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

@@

```{r}
estimate_relation(m_e) |> 
  ggplot(aes(x=Gender, y=Predicted)) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color = Gender), position=position_dodge(width=0.5)) +
 scale_color_manual(values=c("Male"="#E91E63", "Female"="#FF9800"), guide="none") +  labs(y="Enticement") +
  theme_abyss()
```

@@

In general, males tended to assign higher valence ratings to images more frequently. Females, on the other hand, showed higher frequencies of ratings in the middle and lower ranges of the valence scale.

The pattern of valence ratings for images, whether believed or not believed, was similar for both genders, with ratings for believed images closely resembling those for non-believed images.

```{r}
valence <- df |> 
  ggplot(aes(x=Valence, fill = Gender)) +
  geom_bar() +
  scale_y_sqrt(expand = c(0, 0)) +
   scale_x_continuous(
    breaks = c(0, 1),
    labels = c("Not at all","Very much")) +
  scale_fill_manual(values=c("Female"="#E91E63", "Male"="#FF9800"), guide="none") +
  labs(title = "Valence", ) +
    facet_wrap(~Gender) +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

@@

df |> 
  mutate(ConditionBelief = as.factor(ConditionBelief)) |> 
  mutate(ConditionBelief = case_when(
    ConditionBelief == "False" ~ "Not Believed",
    ConditionBelief == "True" ~ "Believed",)) |>
  ggplot(aes(x=Valence, fill=Gender)) +
  geom_bar(position = "identity", binwidth = 0.5) +
  facet_grid(Gender~ConditionBelief, scales="free_y") +
  scale_fill_manual(values=c("Female"="#E91E63", "Male"="#FF9800"), guide="none") +
  # scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Valence to Ratings for Images Based on Belief in Original Labels") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

@@

@@

```{r}
estimate_relation(m_v) |> 
  ggplot(aes(x=Gender, y=Predicted)) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color = Gender), position=position_dodge(width=0.5)) +
 scale_color_manual(values=c("Male"="#E91E63", "Female"="#FF9800"), guide="none") +  labs(y="Valence") +
  theme_abyss()
```

@@

marginaleffects::avg_slopes(m_arousal, variables="Condition", by="Gender") 
# > results_table()

@@

The analysis revealed a significant interaction effect only for **positive attitudes towards AI** among females. Specifically, females with higher positive attitudes towards AI rated AI-generated images as more arousing, relative to when they viewed images labelled as photographs.

@@

m <- glmmTMB::glmmTMB(Arousal ~ Gender / (Condition * SexualActivity_num) + (Condition | Participant) + (1|Stimulus),
                      data=filter(df, ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

@@

This model examines how **Gender and Condition influence enticement** ratings, treating participant and items as random intercepts.

@@

@@

Looking at **feedback as a moderator**, the results showed that enticement ratings for images labelled as AI-Generated were moderated by the feedback that AI is less attractive for males, and that difference between the AI-Generated images and real photos is obvious for females.

@@

           "Feedback_DiffObvious", "Feedback_DiffNone",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing",
           "Feedback_AIMoreAttractive", "Feedback_AILessArousing",
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  moderator_believed <- rbind(moderator_believed,
                             test_moderator(filter(df, ConditionBelief == "True"), "Enticing", m))
}

@@

This model examines the effect of **Gender and Condition on Valence** ratings, while accounting for random variability due to participants and items (i.e., random intercepts).

@@

```{r}
moderator_believed <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "BAIT_ExpertAI",
           "BAIT_Positive", "BAIT_Negative", "PornFrequency", "SexualActivity",

           "Feedback_DiffObvious", "Feedback_DiffNone",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing",
           "Feedback_AIMoreAttractive", "Feedback_AILessArousing",
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  moderator_believed <- rbind(moderator_believed,
                             test_moderator(filter(df, ConditionBelief == "True"), "Valence", m))
}

(filter(moderator_believed, p < .1))
```

@@

```{r}
df$Emotionality <- as.numeric(normalize(predict(pca)$PC1))

emotionality <- df |> 
  mutate(Emotionality_Extreme = ifelse(Emotionality %in% c(0, 1), "Extreme", "Not extreme")) |>
  ggplot(aes(x=Emotionality, fill = Gender)) +
  geom_histogram(aes(color=Emotionality_Extreme), position = "identity", bins=20) +
  facet_grid(~Gender, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Male"="#E91E63", "Female"="#FF9800"), guide="none") +
  scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Emotionality") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

@@

This model looks at the effect of **Gender and Condition on Emotionality**, accounting for random variability due to participants and items (i.e., random intercepts).

@@

```{r}
m <- glmmTMB::glmmTMB(Emotionality ~ Gender / Condition + (1| Participant) + (1|Stimulus),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

@@

Looking at **feedback** as a moderator,  the results show that emotionality ratings for AI-generated images were significantly moderated only by positive attitudes toward AI in females and negative attitudes toward AI in males.

Interestingly, for females, higher positive attitudes (BAIT Positive) were associated with lower emotionality ratings of AI-generated images. For males, higher negative attitudes (BAIT Negative) were associated with higher emotionality ratings.

@@

```{r}
moderator_believed <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "BAIT_ExpertAI",
           "BAIT_Positive", "BAIT_Negative", "PornFrequency", "SexualActivity",

           "Feedback_DiffObvious", "Feedback_DiffNone",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing",
           "Feedback_AIMoreAttractive", "Feedback_AILessArousing",
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  moderator_believed <- rbind(moderator_believed,
                             test_moderator(filter(df, ConditionBelief == "True"), "Emotionality", m))
}

(filter(moderator_believed, p < .1))
```

@@

**Gender Differences**
- Women rate stimuli on average significantly lower in arousal, enticiment and valence than men, however this difference was not significant for valence.

-Males rated AI-generated images significantly lower than photographs in arousal, valence, and enticement. Females showed no significant difference in ratings between AI-generated and photographic images across all outcomes.

**Moderators**
- For women, **positive attitudes towards AI** moderated arousal, enticement, and valence ratings, resulting in **more favorable evaluations of AI-generated images**. Negative attitudes moderated valence ratings, leading to less favorable evaluations.

- For men, negative attitudes towards AI moderated only enticement ratings, with more negative views leading to less enticing ratings of AI-generated images.

- Men’s **perceptions of the attractiveness of AI-generated images** moderated both arousal and enticement ratings. More attractive AI images were rated as more arousing and enticing, while less attractive AI images were rated as less arousing.

- Women’s feedback about the differences between AI and real images being obvious moderated both enticement and valence ratings, with AI-generated images rated as less enticing and less valent.

@@

## Notes/ Questions

1. 'Other' was removed due to the very small percentage of individuals identifying as such.
2. 'Porn' factor was not computed, instead sexual activity and frequency of watching porn were treated seperate.

### 2025-05-08 study2/analysis/3_realness.qmd
data analysis
lacking final figures and rendering emotions analysis script

---
title: "FictionEro2 - Determinants of Reality Beliefs"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

## Data Preparation

@@

@@

Attitudes towards AI did not significantly moderate the influence of Gender and Condition on Realness ratings

@@

<!-- This graph depicts differences in realness ratings based on beliefs about AI's visual abilities for both genders when presented with images labeled as photographs or AI-generated. Notably, images labeled as photographs consistently received higher ratings of realness than those labelled as AI-Generated, even at lower levels of belief in AI's visual capacity.  -->

<!-- As beliefs in AI's ability to generate realistic images increase, both genders show higher realness scores. However, for females with strong beliefs in AI's visual capabilities, realness ratings are higher for images labelled as AI-Generated compared to images labelled as photographs.  -->

<!-- **Notably NONE of these results were significant.** -->

<!-- ```{r} -->
<!-- #| warning: false -->

<!-- m1 <- glmmTMB::glmmTMB(Realness ~ Gender / -->
<!--                         (Condition *  BAIT_Visual) +  -->
<!--                         (1| Participant) + (1|Stimulus), -->
<!--                       data=df,  -->
<!--                       family=glmmTMB::ordbeta(), -->
<!--                       control = glmmTMB::glmmTMBControl(parallel = 8)) -->

<!-- summary(m1) -->

<!-- # stars1 <- data.frame(Gender = as.factor(c("Male", "Male", "Female", "Female")),  -->
<!-- #                     Condition=c("Photograph", "AI-Generated", "Photograph", "AI-Generated"),  -->
<!-- #                     label=c("***", "***", "***", "***"), -->
<!-- #                     x=c(0.6, 0.6, 0.65, 0.65), -->
<!-- #                     y=c(0.8, 0.6, 0.72, 0.55)) -->

<!-- pred1 <- estimate_relation(m1) -->
<!-- p1 <- pred1 |> -->
<!--   ggplot(aes(x=BAIT_Visual, y=Predicted)) + -->
<!--   geom_ribbon(aes(ymin=CI_low, ymax=CI_high, fill=Condition), alpha=0.3) + -->
<!--   geom_line(aes(color=Condition), linewidth=1, key_glyph = draw_key_rect) + -->
<!--   # geom_text(data=stars1, aes(x=x, y=y, label=label, color=Condition), hjust=0.5, size=3) + -->
<!--   # geom_text(data=data.frame(Gender=as.factor("Female"), x=0.65, y=0.638, label="**"), aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=3) + -->
<!--   facet_grid(~Gender) + -->
<!--   scale_y_continuous(labels=scales::percent) + -->
<!--   scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) + -->
<!--   scale_fill_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336"), guide="none") + -->
<!--   theme_minimal() + -->
<!--   theme(axis.text.y = element_text(size = 8), -->
<!--         strip.placement = "outside", -->
<!--         strip.background.x = element_rect(fill=c("lightgrey"), color=NA), -->
<!--         strip.text.x = element_text(size = 10), -->
<!--         strip.text.y = element_text(size = 10), -->
<!--         axis.text.x = element_text(size=9, color="black"), -->
<!--         legend.text = element_text(size = 10)) + -->
<!--   labs(y="Realness\n", fill="Images presented as:", color="Images presented as:", -->
<!--        x = "\nBeliefs about Artificial Information Technology (BAIT)\n'AI algorithms can generate realistic images'", -->
<!--        tag = "A") -->
<!-- p1 -->
<!-- ``` -->

@@

This model evaluates how ratings of realness, predicted by sex, condition, and relevance, are moderated by participants' report of how often they engaged in sexual activity and how often they view pornography.

@@

For males, the belief that AI-generated images were less arousing significantly increased realness ratings for images labeled as AI-generated (compared to photos).

For male participants, the belief that all images were real also significantly increased realness ratings for photos.

For female participants, the belief that no faces were attractive was associated with significantly higher realness ratings for photographs.

@@

stars1 <- data.frame(Gender = as.factor("Male"),
                     Condition = "AI-Generated",
                     label = "***",
                     x = 1.5, 
                     y = 0.55)

@@

p1 <- pred1 |> 
  ggplot(aes(x=Feedback_AILessArousing, y=Predicted)) +
  geom_line(aes(color=Condition, group=Condition), linewidth=.3, position = position_dodge(width=0.2), key_glyph = "path") +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2), size=0.8, linewidth=1, key_glyph = "blank") +
  # geom_text(data=stars2, aes(x=x, y=y, label=label, color=Condition), hjust=0.5, key_glyph = "blank", size=3) +
  geom_text(data=stars2, aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=3) +
  guides(color=guide_legend(override.aes = list(linewidth = 1.5))) +
  facet_grid(~Gender) +
  scale_y_continuous(labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        axis.text.x = element_text(size=9, color="black"),
        legend.text = element_text(size = 10)) +
  labs(y="Realness\n", fill="Images presented as:", color="Images presented as:",
       x = "\n'AI-Images were less arousing' Feedback",
       tag = "B")

@@

>>>>

---
title: "FictionEro2 - Determinants of Reality Beliefs"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---
## Data Preparation

@@

Here, we are examining the effect of various physical characteristics of the images to evaluate whether they influence the perceived realness of the images. Neither Entropy, Complexity, Contrast, nor Luminance significantly influence the perceived realness of the images, nor the likelihood of realness being above 0.5 ("RealnessBelief"), as indicated by the lack of significant effects.

make_model <- function(f, family=glmmTMB::ordbeta()) {
  m <- glmmTMB::glmmTMB(as.formula(f), 
                        data=df, 
                        family=family,
                        control = glmmTMB::glmmTMBControl(parallel = 8))
  print(results_table(m))
  m
}

m <- make_model("Realness ~ Entropy + (1|Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Entropy + (1|Participant) + (1|Stimulus)", family="binomial")
m <- make_model("Realness ~ Complexity + (1|Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Complexity + (1|Participant) + (1|Stimulus)", family="binomial")
m <- make_model("Realness ~ Contrast + (1|Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Contrast + (1|Participant) + (1|Stimulus)", family="binomial")
m <- make_model("Realness ~ Luminance + (1|Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Luminance + (1|Participant) + (1|Stimulus)", family="binomial")
```

@@

This model looks at whether realness ratings are moderated by attitudes towards AI.

Looking at the results, the influence of Gender and Condition on Realness scores was not moderated by attidues towards AI.

@@

@@

This model evaluates how ratings of realness, predicted by Gender and Condition, are moderated by participants' report of how often they engaged in sexual activity and how often they view pornography.

@@

For males, the feedback that AI-generated images were less arousing significantly moderated realness ratings for images labeled as AI-generated, with higher realness ratings for those images when participants thought they were less arousing.

For males, the belief that all images were real significantly increased realness ratings for all images, regardless of their actual content.

For females, the belief that no faces were attractive significantly increased realness ratings for the images, suggesting that lower attractiveness perceptions made the images seem more real.

@@

# stars1 <- data.frame(Gender = as.factor("Male"),
#                      Condition = "AI-Generated",
#                      label = "***",
#                      x = 1.5, 
#                      y = 0.55)
#

@@

p1 <- pred1 |> 
  ggplot(aes(x=Feedback_AILessArousing, y=Predicted)) +
  geom_line(aes(color=Condition, group=Condition), linewidth=.3, position = position_dodge(width=0.2), key_glyph = "path") +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2), size=0.8, linewidth=1, key_glyph = "blank") +
  # geom_text(data=stars2, aes(x=x, y=y, label=label, color=Condition), hjust=0.5, key_glyph = "blank", size=3) +
  # geom_text(data=stars1, aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=3) +
  guides(color=guide_legend(override.aes = list(linewidth = 1.5))) +
  facet_grid(~Gender) +
  scale_y_continuous(labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        axis.text.x = element_text(size=9, color="black"),
        legend.text = element_text(size = 10)) +
  labs(y="Realness\n", fill="Images presented as:", color="Images presented as:",
       x = "\n'AI-Images were less arousing' Feedback",
       tag = "B")

p1

@@

- For males the realness ratings were lower for images labelled AI-Generated compared to real photographs. For females they also rated images labelled as AI-Generated as less Real, altough this effect was not significant. 
- Looking at moderators, attitudes towards AI, participants reports of how often they engaged in sexual activity and how often they view pornography, did not moderate realness ratings
- Looking at the feedback as a moderator for realness ratings, for males, feedback that AI-generated images were less arousing increased realness ratings for those images, and believing all images were real raised realness ratings across all images. For females, the belief that no faces were attractive raised realness ratings for the images.

## Figures

## Notes/ Questions

1. 'Other' was removed due to the very small percentage of individuals identifying as such.
2. 'Porn' factor was not computed, instead sexual activity and frequency of watching porn were treated seperate. 
3. Missing significance starts on figures.
